---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Package for spatial analysis

```{r}
library(sf) 
```

Package's documentation:https://cran.r-project.org/web/packages/sf/index.html

## Other packages

```{r}
library(tidyverse)
```


## Read/Inspect shapefile

```{r}
regions.sf <- 
  sf::read_sf("data/Reg01012020_g_WGS84.shp")
```

```{r}
st_crs(regions.sf)
```

The coordinate reference system is important as it defines the unit (e.g. meters or degrees) for the grid size. In this case the CRS is `WGS 84 / UTM zone 32N` or `ID["EPSG",32632]]` which is in meters (see https://epsg.io/32632).

```{r}
regions.sf %>%
  ggplot2::ggplot() +
  ggplot2::geom_sf()
```

# Union

You want a mask to your grid, so if you gave multiple features (here 20 regions) you might want to combine them (resolving the internal boundaries). With sf and dplyr this can achieved with:

```{r}
regions.sf$unit <- 
  1

italy.sf <- 
  regions.sf %>%
  dplyr::group_by(unit) %>%
  dplyr::summarize(AREA = sum(SHAPE_AREA))
```

```{r}
italy.sf %>%
  ggplot2::ggplot() +
  ggplot2::geom_sf()
```


## Making the grid

This is a two step process. 

1. With `st_make_grid` we create the grid as simple feature object. `cellsize` is in the unit of the mask feature (here `italy.sf` is in meters). `square = FALSE`  creates an hexagonal grid instead of a square grid.

2. With `st_sf` we append to the simple feature object a data frame - for future data analysis. 


```{r}
italy_hex.sf <-
  sf::st_make_grid(italy.sf,
                   cellsize = 15000,
                   what = "polygons",
                   square = FALSE)
```

```{r}
italy_hex.sf <-
  sf::st_sf(hex_id = 1:length(lengths(italy_hex.sf)),
            italy_hex.sf)
```

```{r}
italy_hex.sf %>%
  ggplot2::ggplot() +
  ggplot2::geom_sf()
```


## Cropping

Finally we crop the grid 

```{r}
italy_hex_cropped.sf <-
  sf::st_intersection(italy_hex.sf, 
                  italy.sf %>% st_make_valid())
```

```{r}
italy_hex_cropped.sf %>%
  ggplot2::ggplot() +
  ggplot2::geom_sf()
```

## Intersecting with points

Let's generate a buch of random points within the feature of `regions.sf`. Note: we use `st_buffer` so that some points is going to lay outside of the area we want to interst. 

```{r}
random_pnt.sf <- 
  sf::st_sample(sf::st_buffer(regions.sf, 50000), 
                size = rep(100, nrow(regions.sf)))
```

Our points are in the same CRS of all the other simple features (by design). If this is not the case, you must convert now to a common CRS with 

```{r eval = F}
random_pnt.sf <- 
  sf::st_transform(random_pnt.sf, 32632)
```

and add a data frame to it

```{r}
random_pnt.sf <-
  sf::st_sf(pnt_id = 1:length(lengths(random_pnt.sf)),
            random_pnt.sf)
```


```{r}
random_pnt.sf %>%
  ggplot2::ggplot() +
  ggplot2::geom_sf() +
   ggplot2::geom_sf(data = italy_hex_cropped.sf, fill = NA, alpha = .2)
```

```{r}
point_intersection.list <-
  sf::st_intersects(random_pnt.sf,
                    italy_hex_cropped.sf)
```

```{r}
point_intersection.list
```

Let's make sure we don't lose any point in the conversion from list to vector in case they are out of the intersecting area (here the area of `italy_hex_cropped.sf`). If they are, `st_intersects` intersect will return a vector of length 0 instead of the index of the intersecting feature. Vectors of length 0 in are lost like tears in the rain when we `unlist()`.

```{r}
point_intersection.list[sapply(point_intersection.list, FUN = function(x) length(x) == 0)] <- NA
```

```{r}
point_intersection.list
```

Then we can add the resulting `italy_hex_cropped.sf$hex_id` to `random_pnt.sf`.

```{r}
random_pnt.sf$hex_id <- 
  italy_hex_cropped.sf$hex_id[unlist(point_intersection.list)]
```

```{r}
head(random_pnt.sf)
```


```{r}
random_pnt.sf %>%
  ggplot2::ggplot() +
  ggplot2::geom_sf(aes(colour = hex_id)) +
  scale_colour_viridis_c()
```

As we can see, `hex_id` is assigned from North West to South East. 
